\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Ronny Gey} studied business informatics at the TU-Chemnitz and
then worked as a research assistant at the University of Leipzig and the
FSU Jena. Since 2018 he has been working as a trainee at the University
Library Leipzig and is studying library and information science at the
IBI at the HU Berlin. \url{https://orcid.org/0000-0003-1028-1670}

\textbf{Katrin Leinweber} studied life sciences, ecology and geology,
then shifted towards process automation, data analysis and software
development. She supports and trains researchers and librarians in these
topics since 2017. \url{https://orcid.org/0000-0001-5135-5758}

\textbf{Alexander Struck} has an academic background in LIS \& CS. He
worked for the content industry and does research on citation networks,
research evaluation and research software discovery. Alexander is CIO of
the Cluster of Excellence -- Matters of Activity.
\url{https://orcid.org/0000-0002-1173-9228}

\textbf{Robert Strötgen} is a historian and information scientist and
has developed scientific software at various institutions such as GESIS
and the Georg Eckert Institute. Since 2016 he has been head of the IT
and research support services department at the University Library of
the TU Braunschweig and is now deputy director.
\url{https://orcid.org/0000-0003-3320-5187}

\textbf{Christian Pietsch} is a computational linguist in Bielefeld
University Library's library technology and knowledge management
department. Recent DFG projects he contributed to include CONQUAIRE
(Continuous quality control for research data to ensure reproducibility)
and ORCID DE. Occasionally, he is involved in running BASE (Bielefeld
Academic Search Engine). \url{https://orcid.org/0000-0001-8778-1273}
